Unrecognized States: Theory and Cases

### 2017-07-04 theory.tex
remove table for inside players' payoffs

\noindent Payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\textbf{<<COMP: Place Figure 2 about here>>}
\vspace{2.5pt}

@@

Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded. If both $s$ and $g$ play Status Quo, then the status quo persists. Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.

If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e. the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2). We interpret one player ceding as that player ceding the issue of status (independence vs. reunification) in exchange for some set of (relatively small) payments from the opposing player. Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.

There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously. We denote the payoffs of war as a lottery $\Omega$. This lottery determines whether the secessionists or government wins the war. For simplicity, we assume the outcome of the war is an absorbing state. Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.

For probabilities $p$ of outright victory and $1-p$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p (W_i^t-\zeta_i), \left(1-p\right)( L_i^t-\zeta_i))$. $\Omega^t \equiv (\omega_g^t, \omega_s^t).$ Players are assumed to approach war lotteries as expected values.

@@

In Section~\ref{sec:main}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_c^t < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e. $Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ where $\sigma$ is the reduction due to sanctions).  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

%\item \textit{The secessionists' continuation values from playing \emph{Cede} must be higher than their continuation values from playing \emph{Fight}.}

%\item \textit{The sanctions must overcome the gap between the \emph{Status Quo} and \emph{Cede} continuation values faster than the gap between the \emph{Status Quo} and \emph{Fight} continuation values closes.}

@@

%The situations where condition 2 or 3 fails and the secessionists initiate war follow the same general logic as political science models in the power transitions literature that predict pre-emptive war as the response to the increasing relative capabilities of a rival (Levy 1997; Powell 1999b). The expected payoffs from war are not increasing, but the alternatives are getting worse. The parallels with preemptive war become more direct if we increase the realism of our assumptions and allow sanctions to have a negative effect not only on the economy (the Status Quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_g\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

The proof of Proposition 3 is immediate. Although under the conditions of Proposition 1 (i.e. when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting. The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.

@@

\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}

\newpage
\begin{center}
\includegraphics[width=6.25in]{Table_1_082214.pdf}
\end{center}

>>>>

@@

Only two outcomes of the stage game do not lead to absorbing states. If both $s$ and $g$ play Status Quo, then the status quo persists. Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.

If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e., the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2). We interpret one player ceding as that player ceding the issue of status (independence vs. reunification) in exchange for some set of (relatively small) payments from the opposing player. Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.

There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously. We denote the payoffs of war as a lottery $\Omega$. This lottery determines whether the secessionists or government wins the war. We assume the outcome of the war is an absorbing state and that players approach war lotteries as expected values. Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.

@@

In Section~\ref{sec:main}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo. Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

@@

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities. The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

The proof of Proposition 3 is immediate. Although under the conditions of Proposition 1 (i.e., when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting. The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.

@@
